---
title: "Management research in Africa from editors of the Academy of Management Journal (April 2016)"
author: "Haley Lorenz"
output: pdf_document
---

Conducting a complete simple linear regression analysis to evaluate GDP per capita (x) as a linear predictor of corruption level (y)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("plyr")
```

```{r}
africa <- read.csv("AFRICA.csv", header=TRUE)
```

Variables are included and basic information
```{r}
summary(africa)
```
Level of government corruption is measured on a 100-point scale, where “0” indicates the highest level of corruption and “100” indicates the country is free of corruption and Gross Domestic Product (GDP) is per capita. 

Regression line
```{r}
beta <- lm(Corrupt~ GDP , data = africa)
```

```{r}
plot(africa$GDP, africa$Corrupt, main="GDP Per Capita as a Linear Predictor of Corruption Level", xlab="GDP per capita", ylab="corruption level")
abline(beta,col="blue",lwd=2,lty=2)
```
From the plot we expect a positive relationship between GDP and Corruption Level.

Coefficients and significance
```{r}
summary(beta)
```

```{r}
confint(beta,parm='GDP', level = 0.95)
```

We are 95% confident that the GDP coefficent is between 0.0006626724 and 0.001306386.

Estimating the Corruption when GPD at 25000.
```{r}
predict(beta,data.frame(GDP=25000), interval="prediction")

```

When the GDP is 25000 we predict the corruption level will be between 30.528 and 70.469


From the book we know that the degree of freedom in political rights is measured on a 7-point scale, where “1” indicates the greatest degree and “7” the smallest degree and level of government corruption is measured on a 100-point scale, where “0” indicates the highest level of corruption and “100” indicates the country is free of corruption. 

Regression line
```{r}
beta <- lm(Corrupt~ PolR , data = africa)
```

```{r}
plot(africa$PolR, africa$Corrupt, main="Degree of Freedom in Political Rights as a Linear Predictor of Corruption Level", xlab="degree of freemdom in political rights", ylab="corruption level", cex.main=.99)
abline(beta,col="blue",lwd=2,lty=2)
```
From the plot we expect a negative relationship between degree of freedom and corruption level.

Coefficients and significance
```{r}
summary(beta) 
```
```{r}
confint(beta,parm='PolR', level = 0.95)
```
We are 95% confident that when political freedom decreases corruption decreases by a value between -9.576205 and  -2.917857.

```{r}
predict(beta,data.frame(PolR=5.5), interval="prediction")

```
We are predict that when the degree of freedom of political rights is 5.5 the corruption level will be between 3.619 and 59.791

GDP has a positive relation on corruption level while the degree of fredom of political rights has a negative relation on corruption level. 
