
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("plyr")
library("car")
```

Sales prices of houses in Tampa, Florida.

For 460 houses that were sold in four neighborhoods of 
Tampa, Florida, the sales price (SALES) was recorded in 2017-2018.

We like to model the sales price based on appraised land value (LAND),
appraised improvement value (IMP) and the neighborhoods (NBHD) of the 
houses. The neighborhoods are:

Town & County (TOWN&CTY), Cheval (CHEVAL), Hunter's Green (HUNGREEN), Davis Isles (DAVISISLE)

```{r}
tampa <- read.csv("TAMSALES4.csv", header=TRUE)
```

Basic information
```{r}
summary(tampa)
```
Measurement Note: Prices are given in $1000. 

Plot the sales price versus appraised land and improvement values
```{r}
scatterplotMatrix(~ SALES + LAND + IMP, smooth = FALSE, data = tampa)
```

There is a positive relationship between appraised prices and sale price. We will now further investigate sale price. 
```{r}
boxplot(SALES ~ NBHD, data = tampa, main="Distribution of sales prices by neighborhood")
```

There are clearly difference in the sales price by neighborhood. Davis Isles
seems to be the neighborhood with the highest sales prices, with some exceptional
prices above 2 million dollar. Town & County seems to be the lowest price
neighborhood. The neighborhoods Cheval, Hunter's Green and Town and County
are all grouped together compared with Davis Isles.

We change NBHD variable to a factor and reorder the factor according to the 
book. 

```{r}
tampa$NBHD = as.factor(tampa$NBHD)
tampa$NBHD = factor(tampa$NBHD, levels = c( "TOWN&CTY","CHEVAL", "DAVISISLE", "HUNGREEN"))
tampa$NBHD = revalue(tampa$NBHD, c("CHEVAL"="CH", "DAVISISLE"="DI", "HUNGREEN"="HG", "TOWN&CTY"="TC")) # make names shorter
```

Now we get the summary statistics of the sales by neighborhood:

**Town & Country**:
```{r}
summary(tampa$SALES[tampa$NBHD == "TC"])
```
**Cheval**:
```{r}
summary(tampa$SALES[tampa$NBHD == "CH"])
```
**Davis Isles**:
```{r}
summary(tampa$SALES[tampa$NBHD == "DI"])
```
**Hunter's Green**:
```{r}
summary(tampa$SALES[tampa$NBHD == "HG"])
```



## MODEL 1
Our first model tries to explain the sales price with a linear
model, based only on appraised land and improvement values.

## Find the regression coefficients for a complete quadratic model
```{r}
beta1 = lm(SALES ~ LAND + IMP , data = tampa)
summary(beta1)
```
We see that the model is useful to explain the average sales price,
as the F-statistic has a p-value in the rounding error range.
Both coefficients are highly significant and the model explains
about 96.6% of the squared variability. 

The s value is about $103,500 so we expect that we can predict 
the average sales price by about $207,000. Especially for the 
lower priced neighborhoods that makes the model not useful. 

Both coefficients are positive as we expect. 
For every $1 increase of appraised land value the sales price 
goes up by around $1.60 and for every $1 increase of appraised 
improvement value the sales price goes up by about $1.10.

## MODEL 2
In addition to the appraised land and improvement values
we now include the neighborhood as a qualitative 
variable in our model. The model is still first order
without interaction terms.

```{r}
beta2 = lm(SALES ~ LAND + IMP + NBHD, data = tampa)
summary(beta2)
```

We already suspect that the model is useful, as it includes 
the first model. The $r^2$ is slightly improved
to 96.7%, the adjusted $r^2_a$ very little. 
The s value has decreased to $101,200, but again 
very minor, so it seems that the inclusion
of the linear neighborhood terms has improved 
the model 1 only very modestly. 

## Model 3
Now we include interaction terms between the appraised 
values and the neighborhoods.

```{r}
beta3 = lm(SALES ~ LAND + IMP + NBHD + LAND:NBHD + IMP:NBHD, data = tampa)
summary(beta3)
```

The $r^2$ value has improved very little, but s has decreased
to $96,180 so the predictive values has improved. 
The F-value indicates again the usefulness of the model.

None of the interaction terms seems significant,
but overall they lead to an improvement.

## Model 4
In the final model we add an interaction term between 
land and improvement appraisal and then also the interaction
of this term with the neighborhood.


```{r}
beta4 = lm(SALES ~ LAND + IMP + I(LAND*IMP) + NBHD + LAND:NBHD + IMP:NBHD + I(LAND*IMP):NBHD, data = tampa)
summary(beta4)
```
Again we have minor a minor improvement in the $r^2$ value and
s is decreased by another $100. 

## Comparison of models.
We compare now successively the nested models to see which ones still 
add significantly to the previous model.

```{r}
anova(beta1, beta2)

anova(beta2, beta3)

anova(beta3, beta4)
```

We see that model 2 significantly improves on model 1 and model 3 
significantly improves on model 2. Each the time the p-values
is far below 1%. But model 4 does not seem to add any value
to model 4, the p-value is about 30%. 
So, it seems that model 3 is our "best" model. 

## Result of model 3:

Based on model 3 we get the following prediction equation
for the expected sales value. The neighborhoods are sorted
by increasing median sales prices.

**Town & Country** neighborhood:

E(SALES) = 52 + 1.07$\times$LAND + 0.87$\times$IMP

**Hunter's Green** neighborhood:

E(SALES) = 52 + 1.07$\times$LAND + 0.87$\times$IMP 
              + 10.69
              - 0.49$\times$LAND + 0.34$\times$IMP
              
E(SALES) = 62.69 + 0.58$\times$LAND + 1.21$\times$IMP   

**Cheval** neighborhood:

E(SALES) = 52 + 1.07$\times$LAND + 0.87$\times$IMP 
              + 22.48
              + 0.60$\times$LAND - 0.03$\times$IMP
              
E(SALES) = 74.48 + 1.67$\times$LAND + 0.84$\times$IMP   
        
**Davis Isles** neighborhood:

E(SALES) = 52 + 1.07$\times$LAND + 0.87$\times$IMP 
              - 142.15
              + 0.66$\times$LAND + 0.25$\times$IMP  
              
E(SALES) = -90.15 + 1.73$\times$LAND + 1.12$\times$IMP       

In each neighborhood the regression parameters are different. 

For the most expensive neighborhood the land and 
the improvement coefficient are larger than
1, so the these values seem to be under appraised. 

For the Cheval neighborhood the same is true for the 
appraised land value, but not for the improvement
value. 

For the Hunter's Green neighborhood the appraised
land value seems to play only less of a factor for
the sales price. It has the lowest of all coefficients. 
So, the land value here is over appraised. 
One might have to look into reasons for this,
maybe the neighborhood is older and the 
appraised values are older too. 

## Finally we look at the residuals.

```{r}
plot(tampa$SALES,  beta3$residuals, main="Residual distribution", xlab="Sales price / $1000",
     ylab="Residual / $1000")
```

We see that the residual get larger with larger sales prices, so one of the 
conditions for applying the regression analysis seems not satisfied.
We will learn in chapter 8 how we can address this problem. 



